Attributions for Poverty in Post-Socialist Countries

### 2014-02-14 article2013.Rmd
Kirjoitin kuvioita vielä vähän auki. Tulen huomenna päivällä pariksi tunniksi ja viilaan paperin mallien ja johtopäätösten osalta Moskovan kuntoon

Table 3 shows the proportion of the total number of respondents in a country which chose a particular explanation of poverty. We assumed that citizens in post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations. As expected results show relatively large variation within country groups. For the most of the post-socialist countries, social and individual blame types of explanation, i.e. injustice in a society and laziness of the poor, are the most popular poverty attributions. The highest support for social blame is found in Ukraine, Latvia, Lithuania, Hungary, Russia, Slovenia, Armenia, Moldova and Azerbaijan. In these countries majority of population blames the injustice in a society for poverty. Totally, in 19 of 21 examined post-socialist countries injustice in a society receives the highest support.

@@

Also, in most post-socialist countries the individual blame is the second most popular explanation of poverty. Exceptions here are Slovenia, Estonia and Georgia where citizens endorse societal explanations more likely than individual explanations. In Czech Republic social explanations as well as individual blame explanation receive public support quite evenly. Finally, the clearest exception in post-socialist regime is Uzbekistan where poverty is attributed strongly with internal terms.

@@

Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 2, 3 and 4). Correlation coefficients at country group1 level are shown on top of each regression line.Figure 2 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010). In regard to post-socialist countries there is substantial variation between countries in economic change. The general trend is quite clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.35) and more to individual blame (.48). In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries. However, this is greatly due to economic decline in Great Britain. Hence, without Great Britain we can not see any significant association between economic change and poverty perceptions in Western Europe.

@@

@@

In regard to the level of democracy we can see great differences between and within groups of countries. Overall association is weak (.19 for social blame and -.44 for individual blame), but in CIS countries the higher level of democracy is clearly indicating higher support for social blame and lower support for individual blame type of explanation. In CEE and in Western Europe the variation in the state of democracy is much smaller than in CIS countries. In these countries the support for blame  types of explanations seems to vanish when the level of democracy increases. Consequently, both social and individual fate types of  explanations gain higher support in more democratic societies.

@@

texreg(list(soc.0,soc.1,soc.Cont,soc.4),
        custom.model.names = model.names.multi.A,
        custom.coef.names = coef.names.multi,
        naive = TRUE,
        #use.packages=TRUE,
        #sideways=TRUE,
        scriptsize=TRUE,
        include.pvalues=TRUE,
        caption="Logistic multilevel random intercept model for social blame type of explanation", 
        caption.above=TRUE)
```

@@

Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterly 37 (3), 413–428.

>>>>

Table 2 shows the proportion of the total number of respondents in a country which chose a particular explanation of poverty. We assumed that citizens in post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations. As expected results show relatively large variation within country groups. For the most of the post-socialist countries, social and individual blame types of explanation, i.e. injustice in a society and laziness of the poor, are the most popular poverty attributions. The highest support for social blame is found in Ukraine, Lithuania, Slovenia, Hungary,  Russia and Latvia. In these countries majority of population blames the injustice in a society for poverty. Totally, in 18 of 21 examined post-socialist countries injustice in a society receives the highest support.

@@

Also, in most post-socialist countries the individual blame is the second most popular explanation of poverty. Exceptions here are Ukraine, Slovenia, Estonia and Georgia where citizens endorse societal explanations more likely than individual explanations. In Armenia and Czech Republic social explanations as well as individual blame explanation receive public support quite evenly. Finally, the clearest exception in post-socialist regime is Uzbekistan where poverty is attributed strongly with internal terms.

@@

Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 3, 4 and 5). Correlation coefficients within the post-socialist countries are shown on top of each regression line. Figure 3 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010). In regard to post-socialist countries there is substantial variation between countries in economic change. The trend is very clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.62) and more to individual blame (.71). In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries. However, this is greatly due to economic decline in Great Britain. Hence, without Great Britain we can not see any significant association between economic change and poverty perceptions in Western Europe.

@@

There is also a substantial variation in the level of human development among post-socialist countries. Figure 4 shows that the higher of human development is associated with higher level of social blame (.34). Association holds also in the case of individual blame meaning that the higher the human deveploment is the lower support for support for individual blame explanation (-.48). In this comparison Western countries are better-off in terms of human development and due to substantial variation in dependent variables it is harder to confirm our hypothesis than in the case of post-socialist countries.

In regard to the level of goodness of government we can see major differences are great as expected. Overall association is substantial (.41 for social blame and -.47 for individual blame), and pointing that higher level of democracy is indicating higher support for social blame, whereas lower support for individual blame type of explanation is linked with low figure in good governance indicator. As expected , the variation in the goodness of government in Western Europe is much smaller than in post-socialist countries. In these countries the support for blame types of explanations seems to vanish when the level of democracy increases. Consequently, both social and individual fate types of  explanations gain higher support in more democratic societies.

@@

@@

texreg(list(soc.0,soc.1,soc.Cont,soc.4),
        custom.model.names = model.names.multi.A,
        custom.coef.names = coef.names.multi,
        #naive = TRUE,
        #use.packages=TRUE,
        #sideways=TRUE,
        scriptsize=TRUE,
        include.pvalues=TRUE,
        caption="Logistic multilevel random intercept model for social blame type of explanation", 
        caption.above=TRUE)
```

@@

Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterl
